---
title: "Einzelbaumvolumen_calculate"
format: html
editor: visual
---

## Calculate Einzelbaumvolumen

Classify the detected tree tops in the strata groups and calculate the volumen for each tree

Input: ttops from remote sensing, wefl from FE and volumina from BI calculated in script 2

Output: Einzelbäume with volumina calculated based on their strata

```{r}
#| echo: false
library(sf)
library(here)
library(dplyr)
library(data.table)
library(ggplot2)
```

Load detected tree tops

```{r}
ttops <- st_read(here("input", "ttops_ndom_solling_2023_bigger5m.gpkg"))
```

Classify trees in height groups

```{r}
ttops_height <- ttops %>% mutate(height_group = case_when(
  Z >=2 & Z <12 ~ "jung",
  Z >=12 & Z <21 ~ "gering_baumholz",
  Z >=21 & Z <30 ~ "mittl_baumholz",
  Z >=30 ~ "stark_baumholz"
))
rm(ttops)
```

Load FE data

```{r}
wefl <- st_read(here("input", "nifis_wefl_poly_fe_ma_neuhaus_dassel.gpkg"))
wefl$STICHTAG <- as.character(wefl$STICHTAG)
```

Join welf data (polygon) to ttops (points) in order to have tree species

```{r}
join_ttops_wefl <- st_join(ttops_height, wefl)
```

Classify trees in tree species groups

```{r}
ttops_height_species <- join_ttops_wefl %>%
  mutate(BANR= as.character(BANR)) %>%
  mutate(treespecies_group = case_when(
  substr(join_ttops_wefl$BANR,1L,1L) ==1  ~ "eiche",
  substr(join_ttops_wefl$BANR,1L,1L) ==2 ~ "buche+alh+aln",
  substr(join_ttops_wefl$BANR,1L,1L) ==3 ~ "buche+alh+aln",
  substr(join_ttops_wefl$BANR,1L,1L) ==4 ~ "buche+alh+aln",
  substr(join_ttops_wefl$BANR,1L,1L) ==5 ~ "fichte+tanne",
  substr(join_ttops_wefl$BANR,1L,1L) ==6 ~ "dougl+kiefer+laerche",
  substr(join_ttops_wefl$BANR,1L,1L) ==7 ~ "dougl+kiefer+laerche",
  substr(join_ttops_wefl$BANR,1L,1L) ==8 ~ "dougl+kiefer+laerche"
))
```

Load Einzelbaumvolumen from bi

```{r}
volumina <- fread(here('output', 'statistic_bi_volumen_hoehe_treespecies.csv'), dec=",")
```

Assign Einzelbaumvolumen to each tree depending on its group.

In ttops_height_species_volumina each row is a tree

```{r}
ttops_height_species_volumina <-  ttops_height_species %>% left_join(volumina, by=c("height_group","treespecies_group"))
ttops_height_species_volumina <- ttops_height_species_volumina %>% rename(vol_mean = mean) 
```

```{r}
head(ttops_height_species_volumina)
```

Save point layer with ttops

```{r}
st_write(ttops_height_species_volumina, here("output", "ttops_height_species_volumina.gpkg"))
```
